
---
title: "DATA 607 Week 7 Assignment - Working with XML and JSON in R"
author: "Kalyanaraman Parthasarathy"
output:
  html_document:
    code_folding: show
    depth: 4
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

I have used the following libraries in this assignment to read HTML, XML and JSON format files.

**XML**

**rjson**

All the files are stored in Github repository [DATA 607 Week 7 Assignment - Working with XML and JSON in R](https://github.com/kalyanparthasarathy/DATA607)

**books.html - HTML format file**

**books.xml - XML format file**

**books.json - JSON format file**

Please note that the conversion to Dataframe is performed only for the HTML file however the same type of conversion can be done for XML and JSON formats.

# Load Libraries

```{r}
# Clear the console
cat("\014")

# Check if the package is installed. If not, install the package
if(!require('XML')) {
  install.packages('XML')
  library(XML)
}

# Check if the package is installed. If not, install the package
if(!require('rjson')) {
  install.packages('rjson')
  library(rjson)
}
```


# Read HTML

```{r}
# Read the books.html fle
booksHTMLFile <- readLines("https://raw.githubusercontent.com/kalyanparthasarathy/DATA607/master/books.html")

# Find the lines that matches the pattern "<td> .... </td>"
tdPattern = "<td>([0-9A-Za-z\\., \\(\\)-\\']*)</td>"

# Extract the lines that matches the pattern
matchingLines = grep(tdPattern, booksHTMLFile[1:length(booksHTMLFile)], value=TRUE)

# Extract onlt the information that we need from <td> tag
booksInformation <- trimws(unlist(sub("<td>([0-9A-Za-z\\., \\(\\)-\\']*)</td>", "\\1", matchingLines)))

# Books HTML data as it is - Before formatting
booksInformation
```

## R Structure of HTML

```{r}
str(booksInformation)
```


## Conversion to Dataframe

```{r}
# Books information as data frame
booksInformationDF <- cbind.data.frame(split(booksInformation, rep(1:5, times=length(booksInformation)/5)), stringsAsFactors=F)
names(booksInformationDF) <- c("Title", "Author", "Publication_Year", "Publisher", "ISBN")
booksInformationDF = booksInformationDF[-1,]

# HTML Contents as Dataframe
booksInformationDF

# Structure of HTML Books
str(booksInformationDF)
```


# Read XML

```{r}
# Reading Books XML File
# First download the file from Github repository and save to local file system
# Then use XML parse 
download.file("https://raw.githubusercontent.com/kalyanparthasarathy/DATA607/master/books.xml", destfile = "Books_XML_File.xml")
booksXMLFile <- xmlParse("Books_XML_File.xml")
booksXMLDF <- xmlToDataFrame(booksXMLFile)

# XML Contents
booksXMLDF

```



## R Structure of XML

```{r}
# Structure of XML Books
str(booksXMLDF)

```



# Read JSON

```{r}
# Reading Books JSON File
jsonData <- rjson::fromJSON(file="https://raw.githubusercontent.com/kalyanparthasarathy/DATA607/master/books.json")
```



## R Structure of JSON

```{r}
# Structure of JSON data
str(jsonData)
```



```{r}
```



